Supplementary material

### 2021-06-16 submission/supplement.rmd
create supplement

---
title: "Protocol"
output:
    bookdown::pdf_document2: default
    bookdown::word_document2: default
geometry: margin=1.0in
toc: false
font-size: 12pt
header-includes:
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \newcommand\given[1][]{\:#1\vert\:}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jamia.csl
---

@@

# Introduction
Risk-based HTE analyses typically result in risk stratum specific estimates of treatment effects. Different approaches to individualize treatment effects have been proposed, but it is unclear which approach performs best under different circumstances. We aim  to compare different approaches to  estimating individualized treatment effects.

For all patients we observe covariates $x_1,\dots,x_8$, of which $4$ are continuous and $4$ are binary. More specifically,

@@

# Settings
## Base-case scenario
The base-case scenario assumes a constant odds ratio of $0.8$ in favor of treatment. The simulated datasets are of size $n=4250$, where treatment is allocated at random using a 50/50 split (80% power for the detection of an unadjusted OR of 0.8, assuming an event rate of 20% in the untreated arm). Outcome incidence in the untreated population is set at $20\%$. For the development of the prediction model we use the model defined in \@ref(eq:p0) including a constant treatment effect. When doing predictions, $t_x$ is set to $0$. The value of the true $\bm{\beta}$ is such that the above prediction model has an AUC of $0.75$.

@@

# Risk modeling
Merging treatment arms, we develop prediction models including a constant relative treatment effect:

@@

# Approaches to individualize benefit predictions

## Risk stratification
Derive a prediction model using the same approach as above and divide the population in equally sized risk-based subgroups. Estimate subgroup-specific absolute benefit from the observed absolute differences. Subject-specific benefit predictions are made by attributing to individuals their corresponding subgroup-specific estimate.

## Constant treatment effect
Assuming a constant relative treatment effect, fit the adjusted model in \@ref(eq:risk). Then, an estimate of absolute benefit can be derived from $$\hat{f}_{\text{benefit}}(lp\given \bm{x}, \hat{\bm{\beta}}) = g(lp) - g(lp+\hat{\gamma}) $$

## Linear interaction
The assumption of constant relative treatment effect is relaxed modeling a linear interaction of treatment with the risk linear predictor: 
$$E\{y\given \bm{x}, t_x, \hat{\bm{\beta}}\} = g\big(lp+(\delta_0+\delta_1lp)t_x\big)$$
We predict absolute benefit from $$\hat{f}_{\text{benefit}}(lp\given\bm{x}, \hat{\bm{\beta}}) = g(lp) - g\big(\delta_0+(1+\delta_1)lp\big)$$

## Non-linear interaction
Finally, we drop the linearity assumption and predict absolute benefit by taking the difference between smooth fits, separately derived in each treatment arm:
$$f_{\text{benefit}}(lp\given\bm{x},\hat{\bm{\beta}}) = \hat{f}_{\text{smooth}}(lp\given \bm{x}, \hat{\bm{\beta}}, t_x=0) - \hat{f}_{\text{smooth}}(lp\given \bm{x}, \hat{\bm{\beta}}, t_x=1)$$

@@

# Evaluation
Assuming that $\tau(\bm{x})=E\{y\given\bm{x}, t_x=0\} - E\{y\given\bm{x},t_x=1\}$ is the true benefit for each patient and $\hat{\tau}(x)$ is the estimated benefit from a method under study, the ideal loss function to use for the considered methods would be the unobservable root mean squared error $E\big\{(\hat{\tau} - \tau)^2\given \bm{x}\big\}$. However, in simulations the true patient-level benefit is available. Therefore, we will estimate the RMSE from
$$\text{RMSE}=\frac{1}{n}\sum_{i=1}^n\big(\tau(\bm{x}_i) - \hat{\tau}(\bm{x}_i)\big)^2$$
We also compare the predictive performance of the methods under study. We assess
discrimination using the c-for-benefit statistic described in
[@vanKlaveren2018]. Patients in each treatment arm are ranked based on their
predicted benefit and then are matched 1:1, dropping patients in the larger
treatment arm without a pair. We define *observed* pair-specific treatment
benefit as the difference of observed outcomes between the untreated and the
treated patient of each pair. Pair-specific *predicted* benefit is defined as
the average of predicted benefits within each pair. Then, c-for-benefit is
defined as the probability that from two randomly chosen matched patient pairs
with unequal observed benefit, the pair with greater observed benefit also has a
higher predicted benefit.

We evaluate calibration in a similar manner, using the integrated calibration
index (ICI) for benefit [@Austin2019]. After creating pairs based on predicted
benefit, observed benefits are regressed on the predicted benefits using a
locally weighted scatterplot smoother (loess). The ICI is the area between the
loess fit and the diagonal, with values closer to $0$ indicating better
calibration.

In all cases, the evaluation metrics are calculated in a large simulated
super-population of 500,000 patients.

# References
\setlength{\parindent}{-0.25in}
\setlength{\leftskip}{0.25in}
\noindent
<div id="refs"></div>
\setlength{\parindent}{0in}
\setlength{\leftskip}{0in}
\noindent

\newpage
# Supplement
```{r paok, echo=FALSE, eval=TRUE, warning=FALSE}
table <- read_csv(
  here("data/processed/analysisIds.csv"),
  col_types = "iffiiddddddddddddd"
) %>%
  select(-"type")

>>>>

---
title: "Supplementary material"
output:
    bookdown::pdf_document2: default
geometry: margin=1.0in
toc: false
font-size: 12pt
header-includes:
  - \usepackage{amssymb}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \newcommand\given[1][]{\:#1\vert\:}
editor_options: 
  chunk_output_type: console
bibliography: references.bib
csl: jce.csl
---

@@

# Simulation settings
For all patients we observe covariates $x_1,\dots,x_8$, of which $4$ are
continuous and $4$ are binary. More specifically,

@@

## Base-case scenario
The base-case scenario assumes a constant odds ratio of $0.8$ in favor of treatment. The simulated datasets are of size $n=4250$, where treatment is allocated at random using a 50/50 split (80% power for the detection of an unadjusted OR of 0.8, assuming an event rate of 20% in the untreated arm). Outcome incidence in the untreated population is set at $20\%$. For the development of the prediction model we use the model defined in \@ref(eq:p0) including a constant treatment effect. When doing predictions, $t_x$ is set to $0$. The value of the true $\bm{\beta}$ is such that the above prediction model has an AUC of $0.75$.

@@

## Risk modeling
Merging treatment arms, we develop prediction models including a constant relative treatment effect:

@@

## Approaches to individualize benefit predictions

### Risk stratification
Derive a prediction model using the same approach as above and divide the population in equally sized risk-based subgroups. Estimate subgroup-specific absolute benefit from the observed absolute differences. Subject-specific benefit predictions are made by attributing to individuals their corresponding subgroup-specific estimate.

### Constant treatment effect
Assuming a constant relative treatment effect, fit the adjusted model in \@ref(eq:risk). Then, an estimate of absolute benefit can be derived from $$\hat{f}_{\text{benefit}}(lp\given \bm{x}, \hat{\bm{\beta}}) = g(lp) - g(lp+\hat{\gamma}) $$

### Linear interaction
The assumption of constant relative treatment effect is relaxed modeling a linear interaction of treatment with the risk linear predictor: 
$$E\{y\given \bm{x}, t_x, \hat{\bm{\beta}}\} = g\big(lp+(\delta_0+\delta_1lp)t_x\big)$$
We predict absolute benefit from $$\hat{f}_{\text{benefit}}(lp\given\bm{x}, \hat{\bm{\beta}}) = g(lp) - g\big(\delta_0+(1+\delta_1)lp\big)$$

### Non-linear interaction
Finally, we drop the linearity assumption and predict absolute benefit by taking the difference between smooth fits, separately derived in each treatment arm:
$$f_{\text{benefit}}(lp\given\bm{x},\hat{\bm{\beta}}) = \hat{f}_{\text{smooth}}(lp\given \bm{x}, \hat{\bm{\beta}}, t_x=0) - \hat{f}_{\text{smooth}}(lp\given \bm{x}, \hat{\bm{\beta}}, t_x=1)$$

@@

```{r paok, echo=FALSE, eval=TRUE, warning=FALSE}
table <- read_csv(
  here("data/processed/analysisIds.csv"),
  col_types = "iffiiddddddddddddd"
) %>%
  select(-"type")
